Constrained, yet diverse: Phylogenetic and morphological filtering in epiphytic ferns

### 2019-08-08 ms/manuscript.Rmd
Update MS

@@

Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities (equivalent to the Net Relatedness Index and Nearest Taxon Index, respectively, of @Webb2000 with the sign reversed). Selection of the randomization procedure can strongly bias outcomes [@Gotelli2006;@Ulrich2012]. Our null hypothesis is that fern spores are easily capable of dispersal between terrestrial and epiphytic habitats, and across the relatively short distances separating our plots. We therefore constructed null communities using the "independent swap" algorithm of @Gotelli2000 that maintains species frequency and abundance while randomizing species identities. We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available (`r Ntip(phy)` \si{\species} total; [@Nitta2017]) for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities. We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics. Hereafter, "MPD~phy~" and "MNTD~phy", and "MPD~func~" and "MNTD~func~", will be used to refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.

@@

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife. Raw and processed data are openly available from the Dryad repository at (http://doi.org/[\*\*\* ADD DOI \*\*\*]) [@Nitta2019].

>>>>

Author for correspondence:  
*Joel H. Nitta*  
*Tel: +1 6174960515*  
*Email: joelnitta@gmail.com*

```{r word-count table}
tibble(
  words = c(
    "Total word count (excluding summary, references and legends):_n",
    "Summary:_n",
    "Introduction:_n",
    "Materials and Methods:_n",
    "Results:_n",
    "Discussion:_n",
    "Acknowledgements:_n"),
  figs_tabs = c(
    "No. of figures:_6 (all in color)",
    "No. of tables:_5",
    "No. of supporting information files:_6 (Fig. S1; Table S1-S5)", "", "", "", "")
) %>%
  separate(words, c("words", "count"), sep = "_") %>%
  separate(figs_tabs, c("figs_tabs", "n"), sep = "_", fill = "right") %>%
  mutate(n = replace_na("")) %>%
  gt() %>%
  cols_label(
    words = "",
    count = "",
    figs_tabs = "",
    n = "",
  ) %>% 
  as_rtf() %>%
  write_lines(here::here("ms/word_count.rtf"))
```

@@

Raw values of MPD and MNTD are statistically related to species richness [@Swenson2014]; to obtain metrics that are uncorrelated with species richness, we calculated their standard effect sizes (SES) by comparing observed values to a distribution of \num{999} randomized communities (equivalent to the Net Relatedness Index and Nearest Taxon Index, respectively, of @Webb2000 with the sign reversed). Selection of the randomization procedure can strongly bias outcomes [@Gotelli2006;@Ulrich2012]. Our null hypothesis is that fern spores are easily capable of dispersal between terrestrial and epiphytic habitats, and across the relatively short distances separating our plots. We therefore constructed null communities using the "independent swap" algorithm of @Gotelli2000 that maintains species frequency and abundance while randomizing species identities. We included all species of ferns from Moorea and the neighboring island of Tahiti with genetic data available (`r Ntip(phy)` \si{\species} total; [@Nitta2017]) for phylogenetic null communities, and all species with trait data available (`r fern_traits %>%pull(species) %>% n_distinct`) for functional null communities. We used the \function{ses.mpd} and \function{ses.mntd} functions in the R package \package{picante} [@Kembel2010] to calculate phylogenetic and functional diversity metrics. Hereafter, "MPD~phy~" and "MNTD~phy~", and "MPD~func~" and "MNTD~func~", will be used to refer to the standard effect sizes of MPD and MNTD as measured with phylogenetic and functional trait distances, respectively.

@@

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/canopylife. Raw and processed data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED [@Nitta2019].
